---
output: github_document
---

# Managing data better with pins

Relates to our [discussion on managing and using data](https://github.com/2DegreesInvesting/ds-incubator/issues/35), and extends a previous [introduction to the pins package](https://github.com/2DegreesInvesting/ds-incubator/issues/38).

Here I show how pins meets these requirements:

- [x] Plays well with R.
- [x] Is low cost or better free.
- [x] Data hosted online can also be accessed from a local cache.
- [x] Supports version control with Git.
- [x] Plays well with GitHub.
- [x] Plays well with Azure.
- [x] Allows us to control permissions to read and write data.
- [x] Can handle datasets of the maximum size we need.

tl;dr:

* The pins package meets all requirements with some caveats.
* `pin()`: Pin remote resources locally, work offline and cache results.
* `pin_find()`: Discover new resources across different boards.
* `board_register()`: Share resources in local folders, GitHub, Azure, and more.
* Learn more at <https://pins.rstudio.com/>.

### Plays well with R

It is an [R package](https://cloud.r-project.org/web/packages/pins/index.html), and well integrated with RStudio.

```{r}
# install.packages("pins")
library(pins)
```

Here I also use a few other packages for convenience.

```{r}
library(glue)
library(readr)
library(fs)
library(dplyr, warn.conflicts = FALSE)
```

### Is low cost or better free

As any other R package, pins is free, but [can be used with paid services, e.g. Azure](https://pins.rstudio.com/articles/boards-azure.html).

### Data hosted online can also be accessed from a local cache

Consider a dataset stored online:

* [Nice view](https://github.com/tidyverse/readr/blob/master/inst/extdata/mtcars.csv).
* [Raw view](https://raw.githubusercontent.com/tidyverse/readr/master/inst/extdata/mtcars.csv)

You don't need pins to download or read this file. You may do it with `download.file()` and `read.csv()` (I prefer `readr::read_csv()`). But every time the process may potentially take a long time.

```{r, message=FALSE}
url <- "https://raw.githubusercontent.com/tidyverse/readr/master/inst/extdata/mtcars.csv"
url %>% read_csv()
```

A better way is to `pin()` the URL, which creates a cache in a local "board" that you can later use.

```{r, message=FALSE}
url %>% pin("my_data")

# Works from any R session
my_data <- pin_get("my_data") %>% read_csv()
my_data
```
> `pin()` still works when working offline or when the remote resource becomes unavailable; when this happens, a warning will be triggered but your code will continue to work.

The default location for the local-board cache is convenient and sensible.

```{r}
board_default()

dir_tree(board_cache_path())
```

But you may want to instead use another location, and enable [pins versions](https://pins.rstudio.com/articles/advanced-versions.html).

```{r}
my_cache <- path(tempdir(), "pins_cache")

board_register_local(
  name = "my_local_board", 
  cache = my_cache, 
  versions = TRUE
)

my_data %>% 
  filter(cyl > 6) %>% 
  # Cache new version
  pin(name = "my_data", board = "my_local_board") %>%  
  select(mpg, gear) %>% 
  # Cache new version
  pin(name = "my_data", board = "my_local_board") 

history <- pin_versions("my_data", board = "my_local_board")
history

latest <- history$version[1]
dim(pin_get("my_data", board = "my_local_board", version = latest))

older <- history$version[2]
dim(pin_get("my_data", board = "my_local_board", version = older))
```

In case you are curious, this is the structure of pins versions.

```{r}
dir_tree(my_cache)
```

### Supports version control with Git

For more control, initialize a Git repository in your pins cache.

```{r}
# Set environmental variable so I can use it from the terminal
Sys.setenv(MY_CACHE = my_cache)
Sys.getenv("MY_CACHE")
```

From the terminal:

```{r git, echo=TRUE, engine='/bin/bash'}
git init $MY_CACHE
cd $MY_CACHE
git add .
git commit -m "Start tracking my cache with Git"
```

### Plays well with GitHub

GitHub is a tool we already use a lot and pins plays well with it. 

If you track your cache with Git, you can now push it to GitHub. But GitHub repos can better serve you as a pins board. This works both for public and private repos.

```{r engine="/bin/bash", include=FALSE}
gh repo create maurolepore/demo-board --public
```

For example:

```{r}
board_register_github(
  name = "my_github_board",
  repo = "maurolepore/demo-board",
  cache = my_cache,
  versions = TRUE
)

my_data %>% 
  pin(
    name = "my_data",
    board = "my_github_board",
    description = "Some data",
    repo = "maurolepore/demo-board"
  )

pin_find("my_data", board = "my_github_board")

pin_get("my_data", board = "my_github_board")
```

Worst case scenario you can always get data directly from your local cache.

```{r}
dir_tree(my_cache)

path(my_cache, "my_github_board", "my_data", "data.rds") %>% read_rds()
```

### Plays well with Azure

Azure is important to us because the data managers use it. Data managers may then simply manage access permissions so the right people can access the data they need. Data users like analysts and software developers may then get the access key from Azure, store it safely in their file .Renviron (see [`usethis::edit_r_environ()`](https://usethis.r-lib.org/reference/edit.html)), and read and write data with `pins()`. This way we can use the same familiar `pins()` interface across boards (local, GitHub, Azure, and more) instead of learning a specialized interface (e.g. the [AzureStor](https://cran.r-project.org/web/packages/AzureStor/) package).

Learn more about [Using Azure boards](https://pins.rstudio.com/articles/boards-azure.html).

`board_register_azure()` requires a number of arguments, including secrets, that are best passed as environmental variables -- that is, stored them as key-value pairs in you .Renviron.

```r
# usethis::edit_r_environ()
AZURE_STORAGE_CONTAINER="test-container"
AZURE_STORAGE_ACCOUNT="2diiteststorage"
# Not my real key
AZURE_STORAGE_KEY="ABCABCABCABCABCABCABCABCABCAB=="
```

That setups simplifies how you register an Azure board.

```{r}
board_register_azure()
```

You can now pin resources to the Azure board.

```{r}
datasets::iris %>% 
  pin(name = "iris", board = "azure", description = "My iris data")

datasets::mtcars %>% 
  pin(name = "mtcars", board = "azure", description = "My mtcars data")
```

And you can also find and get resources from the Azure board.

```{r}
pin_find("data", board = "azure")

pin_get("mtcars", board = "azure")
```

And you can also remove resources.

```{r}
pin_remove(name = "mtcars", board = "azure")

pin_find("data", board = "azure")
```


### Allows us to control permissions to read and write data

It is not pins job to control who reads or writes data but it plays well with tools like Azure and GitHub that provide granular [access permissions](https://help.github.com/en/github/getting-started-with-github/access-permissions-on-github).

### Can handle datasets of the maximum size we need

Pins has no limit but some boards do. [Azure](https://pins.rstudio.com/articles/boards-azure.html) is apparently limitless but charges for what it hosts. On [GitHub](https://pins.rstudio.com/articles/boards-github.html) pins can host files up to 2GB in both public and private repos.

--

Cleanup.

```{r, engine="/bin/bash"}
trash $MY_CACHE
```
