---
title: "Using daily data I don't control"
output: github_document
---

```{r}
library(pins)
library(tidyverse)
```

Here is some data I don't control:

```{r}
path_demo_data <- fs::path(tempdir(), "demo-data")
fs::dir_create(path_demo_data)

path_some_data <- fs::path(path_demo_data, "some_data.csv")

"x,y\n1,2\n3,4" %>% 
  read_csv() %>% 
  write_csv(path)

read_csv(path_some_data)
```


```{r}
raw_url <- 
  "https://raw.githubusercontent.com/maurolepore/demo-data/master/some_data.csv"

read_csv(raw_url)
```

It may become unavailable anytime, so I better pin it to a board I do control.

> `pin()` still works when working offline or when the remote resource becomes unavailable; when this happens, a warning will be triggered but your code will continue to work.

-- `?pins::pin()`

I'll register a new GitHub board so I can use Git tools I already know.

1. I'll first create a GitHub repo. Do it however you like; I'll use the terminal.

```bash
# bash
gh repo create --public maurolepore/demo-gh-board
gh repo view maurolepore/demo-gh-board
maurolepore/demo-gh-board
No description provided

No README provided

View this repository on GitHub: https://github.com/maurolepore/demo-gh-board
```

2. I now register that repo as a new pins board.

```{r}
# The default cache is okay, but I include it here just because I can
path_pins_cache <- fs::path(here::here(), "pins_cache")

pins::board_register_github(
  name = "demo-gh-board",
  repo = "maurolepore/demo-gh-board",
  # You should add GITHUB_PAT to .Renviron; see ?usethis::github_token()
  token = usethis::github_token(),  
  cache = path_pins_cache
)

# I shave not created this directory yet
fs::dir_exists(path_pins_cache)
```

I pin the URL to the raw data to my board on GitHub.

```{r}
some_data <- raw_url %>% 
  pins::pin(
    name = "some_data",
    description = "Some raw data",
    board = "demo-gh-board"
  )
```

Although I didn't explicitly read the data, pins stored it in its cache.

```{r}
fs::dir_tree(path_pins_cache)
```

> `pin()` allows you to cache remote resources and intermediate results with ease. When caching remote resources, usually URLs, it will check for HTTP caching headers to avoid re-downloading when the remote result has not changed.

-- `?pins::pin()`

I can later find `some_data` in multiple ways:

* With pins:

```{r}
pin_find("some", board = "demo-gh-board")
```

* With the RStudio addin _Find Pins_:

<img src=http://i.imgur.com/v1nSe43.png width=400 />

* From the tab _Connections_:

<img src=http://i.imgur.com/2KU3b7a.png width=400 />

And I can use the pin anywhere, anytime.

```{r}
path_some_data <- pins::pin_get("some_data", board = "demo-gh-board")
some_data <- read_csv(path_some_data)
```

I would also compute downstream datasets that may be slow to recompute.

```{r}
some_data %>% 
  mutate(new_column = paste(x, y)) %>% 
  pin(
    name = "processed_data", 
    description = "Something expensive to recreate",
    board = "demo-gh-board"
  )
```

What happens if the raw data becomes unavailable?

```{bash engine='/bin/bash'}
# bash terminal
cd ~/git/demo-data
ls
```

```{bash engine='/bin/bash'}
# bash terminal
git rm some_data.csv
git add some_data.csv
git commit -m "Destroy some_data.csv"
git push
```

```{bash engine='/bin/bash'}
# bash terminal
ls
```

No problem, I can still get it.

```{r}
read_csv(pins::pin_get("some_data"))
```

### Cleanup

```{bash engine='/bin/bash'}
# bash terminal
cd ~/git/demo-data
ls
```

```{bash engine='/bin/bash'}
# bash terminal
git revert HEAD -n
git commit -am "Recover some_data"
git push
```

```{bash engine='/bin/bash'}
# bash terminal
ls
```

```{r}
fs::dir_delete(path_pins_cache)
```
